This is a markdown version of the LDA Input Files code.
It 38 samples from 5 different genotypes: ACKG (n=7), ACK (n=8), HACK (n=8), HACKG (n=8) and WT (n=7). 
They can also be combined as ACK/G (n=15), HACK/G (n=16) and WT (n=7)


## Import & Cleanup Normalized counts
```` {r}
library(tibble)
counts <- read.csv("/Users/pwoods/week15August2024/Information/Extra_Files_Needed/Exfoliome-TMM_Normalization.csv")
counts <- counts %>% remove_rownames %>% column_to_rownames(var = "X")
```` 
## Noisy Data Function - Add Noise to Data
This function exists to make the add noise to the data and it helps to prep this early on.
```` {r}
noisy.data <- function(x){
  set.seed(123)
  noise.df <- as.data.frame(matrix(replicate(nrow(x)*ncol(x),rnorm(1,0,0.00001)), ncol = ncol(x), byrow = TRUE))
  data.w.noise <- x + noise.df
  return(data.w.noise)  
}
````

## Import & Cleanup Metadata
```` {r}
library(openxlsx)

meta <- as.data.frame(read.xlsx("/Users/pwoods/week15August2024/Information/Extra_Files_Needed/Mouse_Longitudinal_Study_Metadata-01242023DM.xlsx", sheet = "Joined_Metadata"))
library(dplyr)
meta <- filter(meta,Timepoint == c("W15"))
meta <- meta[,-c(1:6,8:9)]
````

## Order metadata by Genotype
```` {r}
meta <- meta[order(factor(meta$Genotype, levels = c('WT','ACKG','ACK','HACKG','HACK'))),]
````

## Import gene list
```` {r}
colon.genes <- read.xlsx("/Users/pwoods/week15August2024/Information/Extra_Files_Needed/Full_Gene_List-updated-10232023.xlsx")
colon.genes <- filter(colon.genes,Category == c("colon.genes"))
colon.genes <- colon.genes[,-c(1:4,6:11)]
colon.genes <- unique(colon.genes)
````
## Subset counts data
This section contains an additional line to only save lines 1-200; this was an experiment. Do not use this for regular runs.
```` {r}
colon.gene.counts <- counts[rownames(counts) %in% colon.genes,]
colon.gene.counts <- colon.gene.counts[1:200, ] ## adding this - changes this to be 200 genes
````

## Create Key for annotation
```` {r}
numbering <- 0:(nrow(colon.gene.counts)-1)
numbering <- as.data.frame(numbering)
colon.genes <- cbind(numbering,rownames(colon.gene.counts))
names(colon.genes) <- c("Number","Ensembl.ID")
````

## Get names for Mouse Genes for Key
```` {r}
mouse.genes <- read.csv("/Users/pwoods/week15August2024/Information/Extra_Files_Needed/Mouse_biotypes_annotation.csv")[-1]
mouse.genes <- mouse.genes[,-c(3:4)]
names(mouse.genes) <- c("Ensembl.ID","Gene.Name")
colon.genes <- left_join(colon.genes,mouse.genes)
````

## Tidy Data - Colon Genes
```` {r}
colon.genes.num <- paste(nrow(colon.gene.counts))
colon.gene.counts <- as.data.frame(t(colon.gene.counts))
colon.gene.counts <- left_join(meta,rownames_to_column(colon.gene.counts),by=c("Exfoliome.ID"="rowname"))
labels <- colon.gene.counts[,1:3]
````

# Subset Data
This is saved as individual comparisons for the sake of viewing.

## WT.ACK
```` {r}
WT.ACK <- colon.gene.counts %>% filter(Genotype=='WT'|Genotype=='ACK') %>% column_to_rownames(var = "Exfoliome.ID")
WT.ACK <- WT.ACK[,-c(1:2)]
WT.ACK <- noisy.data(WT.ACK)
````

## WT.ACKG
```` {r}
WT.ACKG <- colon.gene.counts %>% filter(Genotype=='WT'|Genotype=='ACKG') %>% column_to_rownames(var = "Exfoliome.ID")
WT.ACKG <- WT.ACKG[,-c(1:2)]
WT.ACKG <- noisy.data(WT.ACKG)
````

## WT.HACK
```` {r}
WT.HACK <- colon.gene.counts %>% filter(Genotype=='WT'|Genotype=='HACK') %>% column_to_rownames(var = "Exfoliome.ID")
WT.HACK <- WT.HACK[,-c(1:2)]
WT.HACK <- noisy.data(WT.HACK)
````

## WT.HACKG
```` {r}
WT.HACKG <- colon.gene.counts %>% filter(Genotype=='WT'|Genotype=='HACKG') %>% column_to_rownames(var = "Exfoliome.ID")
WT.HACKG <- WT.HACKG[,-c(1:2)]
WT.HACKG <- noisy.data(WT.HACKG)
````

## HACK.G.WT
```` {r}
HACK.G.WT <- colon.gene.counts %>% filter(Genotype=='WT'|Grouped.Genotypes=='HACK/G') %>% column_to_rownames(var = "Exfoliome.ID")
HACK.G.WT <- HACK.G.WT[,-c(1:2)]
HACK.G.WT <- noisy.data(HACK.G.WT)
````

## ACK.G.WT 
```` {r}
ACK.G.WT <- colon.gene.counts %>% filter(Genotype=='WT'|Grouped.Genotypes=='ACK/G') %>% column_to_rownames(var = "Exfoliome.ID")
ACK.G.WT <- ACK.G.WT[,-c(1:2)]
ACK.G.WT <- noisy.data(ACK.G.WT)
````

# Write Output files 
This is saved as individual comparisons for the sake of viewing.
In order, each section does the following:
* File name - sets the file name
* Sample count - adds the numbers at the top of the input file indicating the sample number
* Header - saves header and data frame to a file
* Write file - creates file in folder


## File path to folder to save all input files
```` {r}
fp <- c("/Users/pwoods/week15August2024/4-Feature-Testing/")

## Set the count to match the dataset
genes.count <- colon.genes.num
```

## WT.ACK
```` {r}
filename <- paste(fp,"WT_ACK-TestOct2024.txt",sep = "/")
sample.count <- paste(length(which(colon.gene.counts$Genotype=="WT")),",",length(which(colon.gene.counts$Genotype=="ACK")),sep = "")
header <- paste(sample.count,genes.count, sep = "\n")
writeLines(header,filename)
write.table(WT.ACK,filename,append=TRUE,quote=FALSE,row.names = FALSE,col.names = FALSE, eol = "\n")
````

## WT.ACKG
```` {r}
filename <- paste(fp,"WT_ACKG.txt",sep = "/")
sample.count <- paste(length(which(colon.gene.counts$Genotype=='WT')),",",length(which(colon.gene.counts$Genotype=='ACKG')),sep = "")
header <- paste(sample.count,genes.count, sep = "\n")
writeLines(header,filename)
write.table(WT.ACKG,filename,append=TRUE,quote=FALSE,row.names = FALSE,col.names = FALSE, eol = "\n")
````

## WT.HACK
```` {r}
filename <- paste(fp,"WT_HACK.txt",sep = "/")
sample.count <- paste(length(which(colon.gene.counts$Genotype=='WT')),",",length(which(colon.gene.counts$Genotype=='HACK')),sep = "")
header <- paste(sample.count,genes.count, sep = "\n")
writeLines(header,filename)
write.table(WT.HACK,filename,append=TRUE,quote=FALSE,row.names = FALSE,col.names = FALSE, eol = "\n")
````

## WT.HACKG
```` {r}
filename <- paste(fp,"WT_HACKG.txt",sep = "/")
sample.count <- paste(length(which(colon.gene.counts$Genotype=='WT')),",",length(which(colon.gene.counts$Genotype=='HACKG')),sep = "")
header <- paste(sample.count,genes.count, sep = "\n")
writeLines(header,filename)
write.table(WT.HACKG,filename,append=TRUE,quote=FALSE,row.names = FALSE,col.names = FALSE, eol = "\n")
````

## WT.ACK/ACKG
```` {r}
filename <- paste(fp,"WT_ACKG-ACK.txt",sep = "/")
sample.count <- paste(length(which(colon.gene.counts$Genotype=='WT')),",",length(which(colon.gene.counts$Grouped.Genotypes=='ACK/G')),sep = "")
header <- paste(sample.count,genes.count, sep = "\n")
writeLines(header,filename)
write.table(ACK.G.WT,filename,append=TRUE,quote=FALSE,row.names = FALSE,col.names = FALSE, eol = "\n")
````
## WT.HACK/HACKG
```` {r}
filename <- paste(fp,"WT_HACKG-HACK.txt",sep = "/")
sample.count <- paste(length(which(colon.gene.counts$Genotype=='WT')),",",length(which(colon.gene.counts$Grouped.Genotypes=='HACK/G')),sep = "")
header <- paste(sample.count,genes.count, sep = "\n")
writeLines(header,filename)
write.table(HACK.G.WT,filename,append=TRUE,quote=FALSE,row.names = FALSE,col.names = FALSE, eol = "\n")
````
